---
title: "Setup - Rstan"
output: 
  html_document: 
    toc: false
    toc_depth: 3
    toc_float: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing stan and rstan

These instructions are based on:  
https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started and        https://github.com/rmcelreath/rethinking

Go to ``http://mc-stan.org`` and follow the instructions for your platform. The biggest challenge is getting a C++ compiler configured to work with your installation of R. The instructions at the [rstan getting started website](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started) are quite thorough. 

Installing Stan can be a bit complicated on some machines, especially if you haven't run programs that require a C++ compiler before (if that sentance doesn't make sense to you, this means you), so please don't wait until the morning of the workshop to install Stan. Try to install it a few days beforehand, and if you can't get it to work let me know and I'll try to help. 

To check that it works, once you've succesfully done all the other steps please open R and run the code below. If it works, you're good to go!

```{r, echo = TRUE, eval = FALSE}
library(rstan)
stanmodelcode <- "
data {
  int<lower=0> N;
  real y[N];
} 

parameters {
  real mu;
} 

model {
  target += normal_lpdf(mu | 0, 10);
  target += normal_lpdf(y  | mu, 1);
} 
"

y <- rnorm(20) 
dat <- list(N = 20, y = y); 
fit <- stan(model_code = stanmodelcode, model_name = "example", 
            data = dat, iter = 2012, chains = 3, sample_file = 'norm.csv',
            verbose = TRUE) 
print(fit)
traceplot(fit)

```

### `rethinking`

Details on the `rethinking` package:  
https://xcelab.net/rm/software/